### 2022-03-04 paper/localpsf.tex
editing kernel entries section

\subsection{Approximate integral kernel entries}
\label{sec:approximate_kernel_entries}

%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
This section describes how we compute approximate kernel entries
\begin{equation*}
\widetilde{\Phi}(y,x) \approx \Phi(y,x).
\end{equation*}
%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
Let
\begin{equation}
\label{eq:fxyxp}
f_{y,x}(x') := \frac{V(x)}{V(x')}\phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
\end{equation}
It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
$
\Phi(y,x) \approx f_{y,x}(x')
$
when $x'$ is close to $x$. The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly. 
%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
%As discussed in Section SECTION, we make the approximation
%\begin{equation*}
%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
%\end{equation*}
%if $x'$ is close to $x$. Hence, w

We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$. The data used in the interpolation are the following (point,value) pairs
\begin{equation}
\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
%\begin{equation}
%\label{eq:phi_mu_shift32}
%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$. 
%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points.

From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$. But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$. In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$. We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.

\subsubsection{Radial basis function interpolation}

Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
\end{equation}
where $b_i$ are radial basis functions, and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
\begin{equation}
Bw = f,
\end{equation}
where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ and right hand side vector $f \in \mathbb{R}^k$ have entries 
\begin{align*}
B_{ij} :=& b\left(\|x_i - x_j\|\right),\\
f_i :=& f_{y,x}(x_i),
\end{align*}
for $i,j=1,\dots,k$.
% has entries 
%\begin{equation*}
%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
%\end{equation*}
We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
\begin{equation}
b_i(r) := \begin{cases}
r^d, & d=1,3,5,\dots \\
r^d \log r & d=2,4,6,\dots
\end{cases}
\end{equation}
for $r>0$, and $b_i(0):=0$.
%\begin{equation}
%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$. Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.

@@

We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
\begin{equation}
\label{eq:ellipsoid_check93}
(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
\end{equation}
If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation. If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ for such that $x_i \in S_b$, and evaluate
\begin{equation}
f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right).
\end{equation}
%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because 
When the problem is discretized with finite elements, $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we therefore determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell. Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.

\subsubsection{Improved interpolation for symmetric operators}
\label{sec:symmetric_improvements}
If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$. Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$. We could form an alternative approximation of $\Phi(y,x)$ by interpolating the (point, value) pairs
\begin{equation}
\label{eq:phi_mu_shift_y}
\left(y_i,f_{x,y}(y_i)\right), \qquad i=1,\dots,k,
\end{equation}
in the same manner as before, but with $x$ and $y$ now taking opposite roles. We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional. The two interpolation schemes are combined by placing function values 
\begin{equation*}
\left(\widehat{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $x_i-x$, placing function values
\begin{equation*}
\left(\widehat{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.

In detail, define 
\begin{equation*}
\widehat{x_i} := x_i - x \qquad \text{and} \qquad \widehat{y_i} := y_i - y.
\end{equation*}
and let
\begin{equation*}
\widehat{B} = \begin{bmatrix}
\widehat{B}_{xx} & \widehat{B}_{xy} \\
\widehat{B}_{yx} & \widehat{B}_{xx}
\end{bmatrix}, \qquad \widehat{w} = \begin{bmatrix}\widehat{w}_x \\ \widehat{w}_y\end{bmatrix}, \qquad \widehat{f} = \begin{bmatrix}
\widehat{f}_x \\ \widehat{f}_y
\end{bmatrix},
\end{equation*}
be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ be given by
\begin{align*}
\left(\widehat{B}_{xx}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{x}_j\|\right), &\qquad
\left(\widehat{B}_{xy}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{y}_j\|\right), \\
\left(\widehat{B}_{yx}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{x}_j\|\right), &\qquad \left(\widehat{B}_{yy}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{y}_j\|\right),
\end{align*}
for $i,j=1,\dots,k$. We approximate $\Phi(y,x)$ as follows:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widehat{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widehat{y}_i\|\right),
\end{equation}
where $\widehat{w}$ is the solution of the following regularized least-squares problem,
\begin{equation*}
\min_{\widehat{w}} \frac{1}{2}\|\widehat{B} \widehat{w} - \widehat{f}\|^2 + \frac{\alpha}{2}\|\widehat{w}\|^2,
\end{equation*}
for a small regularization parameter $\alpha$. 
We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$. We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are very close to each other, a situation that arises often in practice. Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem. In this case, $\widehat{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard (un-regularized) interpolation equation $\widehat{B} \widehat{w} = \widehat{f}$.

>>>>

\subsection{Approximate integral kernel entries $\widetilde{\Phi}(y,x)$}
\label{sec:approximate_kernel_entries}

%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
%This section describes how we compute approximate kernel entries
%\begin{equation*}
%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
%\end{equation*}
%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
Given $(y,x)\in \Omega \times \Omega$, let
\begin{equation}
\label{eq:fxyxp}
f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined. That is, sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$. In numerical experiments we use $k=10$. We efficiently find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points, and check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem. 
%Per the discussion in Section \ref{eq:local_mean_displacement_invariance}, $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$. The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies widely.

We approximate $\Phi(y,x)$
%compute the approximate kernel entry,
%\begin{equation*}
%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
%\end{equation*}
by interpolating the (point,value) pairs
\begin{equation}
\left(x_i, f_i\right), \qquad i=1,\dots,k
\end{equation}
at the point $x$ using radial basis function interpolation.
%The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
%
%Let
%It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
%and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
%$
%\Phi(y,x) \approx f_{y,x}(x')
%$
%when $x'$ is close to $x$. The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly. 
%%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
%%As discussed in Section SECTION, we make the approximation
%%\begin{equation*}
%%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
%%\end{equation*}
%%if $x'$ is close to $x$. Hence, w
%
%We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$. The data used in the interpolation are the following (point,value) pairs
%\begin{equation}
%\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
%\end{equation}
%and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
%Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
%%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
%%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
%%\begin{equation}
%%\label{eq:phi_mu_shift32}
%%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%%\end{equation}
%%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$. 
%%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
%
%From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$. But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$. In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$. We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
%\subsubsection{Radial basis function interpolation details}
%Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
In detail, we form the approximation $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right).
\end{equation}
Here $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
\begin{equation}
Bw = f,
\end{equation}
where $B \in \mathbb{R}^{k \times k}$ has entries
$
B_{ij} := b\left(\|x_i - x_j\|\right),
$
and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}.
% has entries 
%\begin{equation*}
%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
%\end{equation*}
We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
\begin{equation}
b_i(r) := \begin{cases}
r^d, & d=1,3,5,\dots \\
r^d \log r & d=2,4,6,\dots
\end{cases}
\end{equation}
for $r>0$, and $b_i(0):=0$. For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
%\begin{equation}
%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$. Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.

@@

We evaluate $f_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
\begin{equation}
\label{eq:ellipsoid_check93}
(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
\end{equation}
If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$. If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ such that $x_i \in S_b$, and evaluate
\begin{equation}
f_i = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right),
\end{equation}
as per \eqref{eq:fxyxp} and \eqref{eq:varphi_eval}.
%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because

Note that $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, requires determining which mesh cell contains $y - \mu(x) + \mu(x_i)$, then evaluating the relevant finite element basis functions on that mesh cell. Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.

\subsubsection{Improved interpolation for symmetric operators}
\label{sec:symmetric_improvements}
If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$. Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$. We could form an alternative approximation of $\Phi(y,x)$ by performing the same interpolation as before, but with $x$ and $y$ taking opposite roles. We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional cost. The two interpolation schemes are combined by placing function values 
\begin{equation*}
\left(\widehat{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $x_i-x$, placing function values
\begin{equation*}
\left(\widehat{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
\end{equation*}
at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.

In detail, define 
\begin{equation*}
\widehat{x_i} := x_i - x \qquad \text{and} \qquad \widehat{y_i} := y_i - y.
\end{equation*}
and let
\begin{equation*}
\widehat{B} = \begin{bmatrix}
\widehat{B}_{xx} & \widehat{B}_{xy} \\
\widehat{B}_{yx} & \widehat{B}_{xx}
\end{bmatrix}, \qquad \widehat{w} = \begin{bmatrix}\widehat{w}_x \\ \widehat{w}_y\end{bmatrix}, \qquad \widehat{f} = \begin{bmatrix}
\widehat{f}_x \\ \widehat{f}_y
\end{bmatrix},
\end{equation*}
be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ be given by
\begin{align*}
\left(\widehat{B}_{xx}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{x}_j\|\right), &\qquad
\left(\widehat{B}_{xy}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{y}_j\|\right), \\
\left(\widehat{B}_{yx}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{x}_j\|\right), &\qquad \left(\widehat{B}_{yy}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{y}_j\|\right),
\end{align*}
for $i,j=1,\dots,k$. We approximate $\Phi(y,x)$ as follows:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widehat{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widehat{y}_i\|\right),
\end{equation}
where $\widehat{w}$ is the solution of the following regularized least-squares problem,
\begin{equation*}
\min_{\widehat{w}} \frac{1}{2}\|\widehat{B} \widehat{w} - \widehat{f}\|^2 + \frac{\alpha}{2}\|\widehat{w}\|^2,
\end{equation*}
for a small regularization parameter $\alpha$. 
We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$. We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice. 
%Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem. 
In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.

\subsubsection{Cost to evaluate one approximate kernel entry}

In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system. In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization. In both cases, this requires $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem. Since $k$ is small, this is cheap.
